IML Term project

### 2022-12-05 main_markdown.rmd
Update main_markdown.rmd
Lisäsin osiot, jotka sisältävät VIF-iteraatioiden kautta korreloivien selittävien tekijöiden poistamisen, PCA:n kaikille kolmelle selittävien muuttujien kokoelmille, sekä niihin päivitetyn lasson (testatakseni kuinka paljon malli muuttuu selittävien muuttujien kokoelman muuttuessa). Mitään aiempaa osiota ei ole muokattu, paitsi uuden karsitun selittävien muuttujien arvot on merkitty muiden npf-datojen lisäksi datasettejä kuvaavassa listassa.

# Dataset names:
# npf          - "Original" data: has class2 with values 0,1 as numeric values, 
#                   columns removed: class4, partlybad, date (moved to rowname)
# npf.fact     - Otherwise the same as npf, but class2 is a factor with levels 0 and 1
# npf.168      - Processed data: Parameters that have different measurement heights have been reduced to height 16.8m
#                   columns removed: Parameters with height that's not 16.8m
# npf.168.fact -  Otherwise the same as npf.168, but class2 is a factor

@@

>>>>

# Dataset names:
# npf          - "Original" data: has class2 with values 0,1 as numeric values, 
#                   columns removed: class4, partlybad, date (moved to rowname)
# npf.fact     - Otherwise the same as npf, but class2 is a factor with levels 0 and 1
# npf.168      - Processed data: Parameters that have different measurement heights have been reduced to height 16.8m
#                   columns removed: Parameters with height that's not 16.8m
# npf.168.fact -  Otherwise the same as npf.168, but class2 is a factor
# npfsp        - Reduced set of explanatory variables, received through running VIF analysis and repeating variable with highest VIF until no variable has a VIF
#                 higher than 10
# npfsp.fact   - Otherwise the same as npfsp but class2 is a factor

@@

### vif
```{r vif}
testmod<-glm(class2~., npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean-CS.mean-CS.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean-CS.mean-CS.std-NOx168.mean-NOx168.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

#Thus, we come up to the following collection of variables
coefficients(testmod)
colnames(npf)
npfsp<-npf.168[,c(1:2,5:6,9:10,13:16,19:20,23:24,27:30,37)]
colnames(npfsp)
npfsp.fact<-npfsp
npfsp.fact[,19]<-as.factor(npfsp.fact[,19)
```

### PCA
#PCA for all three versions of data 
```{r PCA}
pcaOrig<-prcomp(npfOrig[,5:104],center = T, scale.=T)
cumsum(pcaOrig$sdev^2)/sum(pcaOrig$sdev^2)
x1<-as.matrix(npfOrig[,5:104])%*%pcaOrig$rotation[,1:12]
pca.168<-prcomp(npf.168[,1:36],center = T, scale.=T)
cumsum(pca.168$sdev^2)/sum(pca.168$sdev^2)
x2<-as.matrix(npf.168[,1:36])%*%pca.168$rotation[,1:12]
pca.sp<-prcomp(npfsp[,1:18],center = T, scale.=T)
cumsum(pca.sp$sdev^2)/sum(pca.sp$sdev^2)
x3<-as.matrix(npfsp[,1:18])%*%pca.sp$rotation[,1:11]
```

### Lasso for reduced sets of variables
```{r lasso2}
lasso2<-cv.glmnet(x=as.matrix(npf.168[,1:36]),y=class2, alpha=1,family="binomial", type.measure="class")
lasso3<-cv.glmnet(x=as.matrix(npfsp[,1:18]),y=class2, alpha=1,family="binomial", type.measure="class")
lassopc1<-cv.glmnet(x=x1,y=class2, alpha=1,family="binomial", type.measure="class")
lassopc2<-cv.glmnet(x=x2,y=class2, alpha=1,family="binomial", type.measure="class")
lassopc3<-cv.glmnet(x=x3,y=class2, alpha=1,family="binomial", type.measure="class")
plot(lasso2)
plot(lasso3)
plot(lassopc1)
plot(lassopc2)
plot(lassopc3)
```

\newpage
